Pipeline Performance Figure Generation

### 2023-05-27 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
Updated Figure 3A.

```{r}
viruses_dw$environment <- "drinking_water"

viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"

@@

recoding AJ's samples to be able to separate virome from metagenome (and remove others)
```{r}
viruses_all$environment2 <- viruses_all$environment
viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake_Erie_virome"
viruses_all$environment2[viruses_all$assembly %in% c("49613", "49614", "49618", "49621",
                                                     "49624", "49628", "49632", "49638")] <- "Lake_Erie_metagenome"
viruses_all <- viruses_all[viruses_all$environment2!="Lake_Erie",]
```

```{r}
viruses_all$environment2 <- factor(viruses_all$environment2,
                                    levels = c("drinking_water", "Lake_Erie_metagenome", 
                                               "global_oceans",
                                               "wastewater", "Lake_Erie_virome", 
                                               "Lake_Michigan"))
```

```{r}
viruses_all$best_coloring <- "others"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 1 1 1 1"] <- "all"
viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 1 1 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 1 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 0 1"] <- "VS2 combo"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 0"] <- "high precision"
```

```{r}
ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)
```

```{r}
viruses_all$type <- "metagenome"
viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake_Erie_virome",
                                                 "Lake_Michigan")] <- "virome"
```

```{r}
ggplot(viruses_all[viruses_all$best_coloring %in% c("high MCC", "high recall"),], aes(x=environment2, y=prop_vir, 
                                  color=type, fill=type)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Environment") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(magma(4)[2:3], 0.3)) +
  scale_color_manual(name="",
                     values = alpha(magma(4)[2:3], 0.7)) +
  facet_wrap(~best_coloring)
```

```{r}
imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            )
```

```{r}
ggplot(imp_toolcombs, aes(x=best_coloring, y=mean, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_point(size=3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment)
```

```{r}
imp_toolcombs_wide <- viruses_all %>% group_by(best_coloring, environment) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            ) %>%
  select(mean, environment, best_coloring) %>%
  pivot_wider(names_from=environment, values_from=mean)
```

>>>>

```{r}
viruses_dw$environment <- "drinking water"

viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake Erie"
viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake Michigan"
viruses_rest$environment[viruses_rest$environment=="tara"] <- "global oceans"

@@

recoding AJ's samples to be able to separate virome from metagenome (and remove others)
```{r}
viruses_all$environment2 <- viruses_all$environment
viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake Erie virome"
viruses_all$environment2[viruses_all$assembly %in% c("49613", "49614", "49618", "49621",
                                                     "49624", "49628", "49632", "49638")] <- "Lake Erie metagenome"
viruses_all <- viruses_all[viruses_all$environment2!="Lake Erie",]
```

```{r}
viruses_all$environment2 <- factor(viruses_all$environment2,
                                    levels = c("drinking water", "Lake Erie metagenome", 
                                               "global oceans",
                                               "wastewater", "Lake Erie virome", 
                                               "Lake Michigan"))
```

```{r}
viruses_all$ruletype <- "other"
viruses_all$ruletype[viruses_all$toolcombo %in% c("0 0 1 0 0 1",
                                                          "0 0 1 1 0 0",
                                                          "0 0 0 1 0 0")] <- "high precision"
viruses_all$ruletype[viruses_all$toolcombo %in% c("1 1 0 0 1 1",
                                                          "1 1 0 1 1 1",
                                                          "1 1 0 0 0 1",
                                                          "1 1 0 1 0 1")] <- "high recall"
viruses_all$ruletype[viruses_all$toolcombo %in% c("0 0 1 0 0 1",
                                                          "0 1 1 0 0 1",
                                                          "0 0 0 0 0 1",
                                                          "1 0 1 1 0 1",
                                                          "1 0 1 0 1 1",
                                                          "1 0 1 0 0 1")] <- "high MCC"
viruses_all$ruletype[viruses_all$toolcombo %in% c("1 1 1 1 1 1")] <- "all"
```

```{r}
viruses_all$type <- "metagenome"
viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake Erie virome",
                                                 "Lake Michigan")] <- "virome"
```

```{r}
ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
    strip.background = element_rect(fill="white", color="grey"),
    strip.text = element_text(color="black")
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
  facet_grid(~environment2, scales = "free")
```

```{r}
viruses_all %>% 
  select(prop_vir, toolcombo, environment2) %>%
  filter(toolcombo=="1 0 0 1 0 0") %>%
  filter(environment2=="wastewater") 
  group_by(environment2) %>%
  summarise(mean = mean(prop_vir))

viruses_all %>% 
  select(prop_vir, toolcombo, environment2) %>%
  filter(toolcombo=="0 1 1 0 1 1") %>%
  group_by(environment2) %>%
  summarise(mean = mean(prop_vir))

viruses_all %>% 
  select(prop_vir, toolcombo, environment2) %>%
  filter(toolcombo=="1 0 0 0 0 1") %>%
  group_by(environment2) %>%
  summarise(mean = mean(prop_vir))

viruses_all %>% 
  select(prop_vir, toolcombo, environment2) %>%
  filter(toolcombo=="1 1 1 1 1 1") %>%
  group_by(environment2) %>%
  summarise(mean = mean(prop_vir))
```

```{r}
fig <- ggplot(viruses_all, aes(x=ruletype, y=prop_vir, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=10),
    axis.text.x=element_text(size=12, angle = 90, vjust = -0.1),
    legend.text=element_text(size=12),
    axis.title=element_text(size=12),
    strip.background = element_rect(fill="white", color="grey"),
    strip.text = element_text(color="black", size=12)
  ) +
  xlab("Environment") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
  facet_wrap(~environment2)

fig

ggsave(
  "../IntermediaryFiles/simplified_environmental.png",
  plot = fig,
  scale = 1,
  width = 6.5,
  height = 4,
  units = c("in"),
  dpi = 300,
  limitsize = TRUE
)
```

```{r}
imp_toolcombs <- viruses_all %>% group_by(ruletype, environment2) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            )
```

```{r}
ggplot(imp_toolcombs, aes(x=ruletype, y=mean, 
                                  color=ruletype, fill=ruletype)) +
  geom_point(size=3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)
```

### 2023-05-27 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Updated Figure 3A.

viruses_long_scores$rule_type <- "viral sorting tool"
viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_tv"] <- "tuning"
viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_ntv"] <- "tuning"

@@

```{r}
pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

>>>>

viruses_long_scores$rule_type <- "viral identification tool-based"
viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_tv"] <- "tuning"
viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_ntv"] <- "tuning"

@@

viruses_long_scores$seqtype <- factor(viruses_long_scores$seqtype, levels = c("virus", "bacteria", "archaea", "plasmid", "protist", "fungi"))

#viruses_long_scores$rule_combination <- factor(viruses_long_scores$rule_combination, 
#                                               levels = c(rev("keep_score_ind_vs2", "keep_score_ind_vs",
#                                                              "keep_score_ind_vb", "keep_score_ind_dvf"),
#                                                         rev("keep_score_ind_tv", "keep_score_ind_ntv")))

viruses_long_scores$rule_type <- factor(viruses_long_scores$rule_type, levels = c("viral identification tool-based", "tuning"))

pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

### 2023-05-27 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated Figure 3A.

  auc <- round(auc(truepositive, keep_score),4)

  performance <- c(precision, recall, F1, MCC, auc)
  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC")

@@

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)),
                       MCC=rep(0, nrow(combos)),
                       AUC=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                            include_virsorter = combos$VS[i],
                                            include_virsorter2 = combos$VS2[i],
                                            include_tuning_viral = combos$tune_viral[i],
                                            include_tuning_not_viral = combos$tune_not_viral[i],
                                            include_deepvirfinder = combos$DVF[i])

@@

## Calculate the performance of each pipeline
quantifying variability between rule sets
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
                      precision=rep(0, nrow(combos_list)*8),
                       recall=rep(0, nrow(combos_list)*8),
                       F1=rep(0, nrow(combos_list)*8),
                       MCC=rep(0, nrow(combos_list)*8), 
                      AUC=rep(0, nrow(combos_list)*8))

@@

```{r}
accuracy_scores_melt <- accuracy_scores %>% 
  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo, ruletype) %>%
  pivot_longer(cols=c(precision, recall, MCC), 
               names_to="performance_metric",
               values_to="performance_metric_score")
```

>>>>

  prop_viral <- (TP+FP)/(TP+FP+TN+FN)

  performance <- c(precision, recall, F1, MCC, prop_viral)
  names(performance) <- c("precision", "recall", "F1", "MCC", "prop_viral")

@@

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)),
                       MCC=rep(0, nrow(combos)),
                       prop_viral=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                            include_virsorter = combos$VS[i],
                                            include_virsorter2 = combos$VS2[i],
                                            include_tuning_viral = combos$tune_viral[i],
                                            include_tuning_not_viral = combos$tune_not_viral[i],
                                            include_deepvirfinder = combos$DVF[i])

@@

## Calculate the performance of each pipeline
quantifying variability between rule sets
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
                      precision=rep(0, nrow(combos_list)*8),
                       recall=rep(0, nrow(combos_list)*8),
                       F1=rep(0, nrow(combos_list)*8),
                       MCC=rep(0, nrow(combos_list)*8), 
                      prop_viral=rep(0, nrow(combos_list)*8))

@@

```{r}
accuracy_scores_melt <- accuracy_scores %>% 
  select(testing_set_index, precision, recall, MCC, prop_viral, numrules, numtools, toolcombo, ruletype) %>%
  pivot_longer(cols=c(precision, recall, MCC, prop_viral), 
               names_to="performance_metric",
               values_to="performance_metric_score")
```

```{r}
fig <- ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="prop_viral",], 
              aes(x=ruletype, y=performance_metric_score, 
                                  color=ruletype, fill=ruletype)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=10),
    axis.text.x=element_text(size=12, angle = 90, vjust = -0.1),
    legend.text=element_text(size=12),
    axis.title=element_text(size=12),
    strip.background = element_rect(fill="white", color="grey"),
    strip.text = element_text(color="black", size=12)
  ) +
  xlab("Environment") +
  ylab("Proportion Called Viral") +
  scale_fill_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
  scale_color_manual(name="",
                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
  geom_hline(yintercept = 0.1, linetype="dashed", color="dark grey")

```{r}
ggsave(
  "../IntermediaryFiles/simplified_performance_metricrule_types.png",
  plot = fig,
  scale = 1,
  width = 6.5,
  height = 4,
  units = c("in"),
  dpi = 300,
  limitsize = TRUE
)
```
